A Convolutional Neural Network for Outsole Recognition\tnoteref{t1}

### 2020-04-14 paper/paper-revision.Rnw
Slowly working my way through this...

\begin{abstract}
\mt{To assess random match probability at the level of class characteristics, one could imagine sampling shoes from a local population and calculating the proportion of shoes "similar" to a given print. Our proposed implemention of such a procedure involves 1) a scanner to gather images of shoes from a local population and 2) software to automatically label images within a feature space and assess the images for similarity among those features.} In this paper, we discuss an algorithm for automatically labeling shoe outsole images using a \mt{classification scheme} similar to what examiners use: a set of features based on geometric shapes such as circles, stars, and quadrilaterals. The automatic labeling algorithm performs well on images with unambiguous shapes and moderate color contrast, though some systematic errors arise from the physical similarities of geometric shapes. Additional improvements may be realized by modifying the labeling scheme to reduce feature overlap and by preprocessing the images to improve contrast. \mt{This new approach represents a significant step toward making tractable the computation of random match probabilities of tread patterns, and could potentially be modified to automate other types of footwear examinations based on class characteristics.}
\end{abstract}

@@

\mt{In the United States, forensic analysis of shoe print evidence is varied and difficult. The process of collecting print evidence from a crime scene is not standardized, and it is often impossible to conduct an individualizing analysis (i.e., one based on randomly acquired characteristics) when a print is partial or of low quality. In fact, the majority of footwear examinations in the United States reach conclusions regarding the correspondence of class characteristics rather than of individualizing characteristics \citep{smith_2011}. In the cases where individualizing characteristics are not used, it may be both possible and useful to assess the random match probability of the footwear evidence at the level of class characteristics; in other words, given a questioned shoe print, what is the probability that a randomly selected shoe in the relevant population has a similar tread pattern? Note that this question is distinct from make and model determination since many shoes of different makes and models can still have similar tread patterns (e.g., knockoffs, similar models over different years, etc).} \fix{[This is an important point. Should I make a bigger deal out of it throughout the paper?]}

>>>>

\begin{abstract}
\svp{Quantitative assessment of class characteristic frequency is a difficult problem, in part because there is currently no way to collect and efficiently process and categorize the quantity of data which would be required for such assessments.
The human labor required to both collect and systematically classify images collected from the population in sufficient quantity to make such assessments has been prohibitive. % I'm not thrilled with this bit, but I do think we need to drive the point home a bit harder as to why we have to automate both parts of this process
In this paper, we discuss a computational method which complements a passive, unmanned scanning device that is currently under development. The automatic class characteristic recognition algorithm described in this paper has been developed with photographs of outsole images, and identifies high-level features} such as circles, stars, and quadrilaterals. \svp{These features are designed to mimic high-level features used by examiners, to facilitate automatic processing of features without resorting to black-box feature sets commonly used in neural networks. The algorithm}  performs well on images with unambiguous shapes and moderate color contrast, though some systematic errors arise from the physical similarities of geometric shapes. Additional improvements may be realized by modifying the labeling scheme to reduce feature overlap and by preprocessing the images to improve contrast. This new approach represents a significant step toward \svp{enabling the collection and application of population-level class characteristic data to support quantitative assessments of the value of class-characteristic based similarity assessments.} % We do NOT want to even hint at automating away anyone's job.
\end{abstract}

@@

% \mt{In the United States, forensic analysis of shoe print evidence is varied and difficult.}% this is very imprecise
\svp{Collection and use of footwear evidence varies heavily across jurisdictions within the United States, and, though it is more common for such analyses to be used in Europe, the U.K., and Israel, many of the same fundamental issues remain.}
\svp{It is often difficult to recover footwear evidence with sufficient detail to identify individualizing characteristics, either because of degradation or because the medium did not record the individualizing characteristics in sufficient detail to examine. It is common for prints to be only partially recorded or recovered, which further compounds the difficulty of performing these types of analysis.}
\svp{As a result, }
% \mt{The process of collecting print evidence from a crime scene is not standardized, and it is often impossible to conduct an individualizing analysis (i.e., one based on randomly acquired characteristics) when a print is partial or of low quality. In fact,
\mt{ the majority of footwear examinations in the United States reach conclusions regarding the correspondence of class characteristics rather than of individualizing characteristics \citep{smith_2011}.
\svp{This tendency is also reflected in the SWGTREAD range of conclusions scale, which has two separate categories to indicate degrees of class characteristic based association out of a range of 7 conclusions.}
\svp{Currently, the guidelines also caution against any statistical assessment because accurate data and models to support statistical calculations do not exist.}
\svp{If it were possible to collect a sufficiently large quantity of footwear impression data from a local population, it would be helpful to be able to calculate the frequency of shoes with similar class characteristics in the population in order to characterize the determinative value of a particular piece of evidence.}
% In the cases where individualizing characteristics are not used, it may be both possible and useful to assess the random match probability of the footwear evidence at the level of class characteristics; in other words, given a questioned shoe print, what is the probability that a randomly selected shoe in the relevant population has a similar tread pattern?
\mt{Note that this question is distinct from make and model determination since many shoes of different makes and models can still have similar tread patterns} \svp{on the class characteristic level.} %\fix{[This is an important point. Should I make a bigger deal out of it throughout the paper?]}% if it fits in, that's a good idea.
